---
title: "Lab 1"
author: "Justin Kleiber"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    number_sections: no
  pdf_document:
    fig_caption: yes
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
Working directory
```{r}
getwd()
```

# Task 2
```{r}
ddt=read.csv("DDT.csv") 
head(ddt)
```

# Task 3
* **Qualitative Variables:** RIVER, SPECIES
* **Quantitative Variables:** MILE, LENGTH, WEIGHT, DDT
* There are three different SPECIES in the ddt data set, as shown below:
```{r}
summary(ddt$SPECIES)
```
* Subsetting the data such that WEIGHT > 800 and SPECIES is LMBASS
```{r}
with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS", ])
```
* Subsetting the data such that RIVER is SCM and DDT > 4.0
```{r}
with(ddt, ddt[RIVER == "SCM" & DDT > 4.0, ])
```

## Clicker Questions
1. 42.8125
```{r}
mean(ddt$LENGTH)
```
2. 376.5461
```{r}
sd(ddt$WEIGHT)
```
3. 0
4. 1.00

# Task 4
Table of the rivers:
```{r}
table(ddt$RIVER)
```

Color coded barplot of the rivers:
```{r}
barplot(table(ddt$RIVER), col = rainbow(4))
```

Table of river name crossed with fish species:
```{r}
table(ddt$RIVER, ddt$SPECIES)
```

barplot of the rivers crossed with the species of fish:
```{r}
barplot(table(ddt$RIVER, ddt$SPECIES))
```

# Task 5
Pie chart of fish species
```{r}
pie(table(ddt$SPECIES))
```

Pie chart of the rivers
```{r}
pie(table(ddt$RIVER))
```

# Task 6
boxplot of DDT
```{r}
with(ddt, boxplot(DDT, ylab="DDT", col="Red", notch=TRUE))
```

boxplot of WEIGHT
```{r}
with(ddt, boxplot(WEIGHT,ylab="WEIGHT",col="Green",notch=TRUE))
```

boxplot of LENGTH
```{r}
with(ddt, boxplot(LENGTH,ylab="LENGTH",col="Blue",notch=TRUE))
```


# Task 7
coplot of LENGTH vs WIDTH given the RIVER
```{r}
river_color = with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Blue",
                               ifelse(RIVER=="SCM","Green","Black"))))
coplot(LENGTH~WEIGHT|RIVER, data = ddt, col = river_color)
```

coplot of DDT vs WEIGHT given the SPECIES
```{r}
fish_color = with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot(DDT~WEIGHT|SPECIES, data = ddt, col = fish_color)
```


# Task 8
GGPlot Boxplot
```{r}
library(ggplot2)
ggplot(ddt, aes(x = SPECIES, y = WEIGHT, fill = RIVER)) + geom_boxplot() + ggtitle("Justin Kleiber")
```

GGPlot Violin Plots
```{r}
ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) + geom_violin() + ggtitle("Justin Kleiber")
```

GGPlot Scatter 
```{r}
ggplot(ddt, aes(x = WEIGHT, y = LENGTH, color = SPECIES)) + geom_point() + ggtitle("Justin Kleiber")
```