---
title: "Lab 2"
author: "Justin Kleiber"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
```{r}
getwd()
```

# Task 2
```{r}
mpg_df = read.csv("EPAGAS.csv")
head(mpg_df)
```

# Task 3
Find the mean and variance of z values
```{r}
x = mpg_df$MPG
z = (x - mean(x)) / sd(x)

mean(z)
var(z)
```

Find the outlier:
```{r}
x[abs(z) > 3]
```

Lattice Plot:
```{r}
library(lattice)

# Color vector
outlier_colors = ifelse(abs(z)>3, "Red",
                 ifelse(abs(z)>=2 & abs(z)<=3,"Blue", "Black"))  


dotplot(x, col = outlier_colors)
```

# Task 4
```{r}
with(mpg_df, boxplot(MPG, main = "Miles Per Gallon EPA Data", col = "Black", notch = TRUE, horizontal = TRUE))
```

Chebyshev says the proportion within 2 standard deviations of the mean will be **at least** 1 - 1/K^2. For K = 2, this yields 3/4, or 75%

Actual proportion:
```{r}
n = length(x)
interval = x[abs(z) <= 2]
ratio = length(interval) / n
ratio
```
96% agrees with the Chebyshev theorem, as it is more than 75%.

The empirical rule says that **approximately** 95% of the distribution should be within two standard deviations of the mean. This corresponds very well - it is only off by 1%! The Empirical rule is valid in this case because, as shown by the boxplot, the distribution is unimodal and symmetrical. This fulfills the assumptions made by the empirical rule.
